---
title: "SQL Sample Code"
author: "Erica Wang"
output: pdf_document
---

## Problem 1. What is customer Karen Ng's address?
```sql
SELECT Street, City, State_Prov, Zip_Postal_Code, 
Country FROM DIVECUST WHERE DIVECUST.`Name`='Karen Ng'
```
#### Output
2134 Elmhill Pike	Klamath Falls	OR	97603	U.S.A.

## Problem 2. At what destination and site(s) might you find a Spotted Eagle Ray?
```sql
SELECT Destination_Name, Site_Name FROM 
BIOLIFE INNER JOIN BIOSITE ON 
BIOLIFE.`Species_No`=BIOSITE.`Species_No` 
INNER JOIN SITES on BIOSITE.`Site_No`=SITES.`Site_No` 
INNER JOIN DEST ON SITES.`Destination_No`=DEST.`Destination_No` 
WHERE BIOLIFE.`Common_Name`='Spotted Eagle Ray'
```
#### Output
```{r,echo=FALSE}
table <- read.csv("query_result.csv")
table
```


## Problem 3. What sites might Lorraine Vega dive on her trip?
```sql
SELECT Site_Name FROM DIVECUST 
INNER JOIN DIVEORDS ON 
DIVECUST.`Customer_No`=DIVEORDS.`Customer_No` 
INNER JOIN DEST ON 
DEST.`Destination_Name`=DIVEORDS.`Destination` 
INNER JOIN SITES ON 
DEST.`Destination_No`=SITES.`Destination_No` 
WHERE DIVECUST.`Name`='Lorraine Vega'
```
#### Output
Palancar Reef, Santa Rosa Reef, Chancanab Reef, Punta Sur, Yocab Reef

## Problem 4. What is the cost of the equipment rental for Louis Jazdzewski?
```sql
SELECT SUM(Rental_Price * Qty) FROM DIVESTOK 
INNER JOIN DIVEITEM ON DIVESTOK.`Item_No`=DIVEITEM.`Item_No` 
INNER JOIN DIVEORDS ON DIVEORDS.`Order_No`=DIVEITEM.`Order_No` 
INNER JOIN DIVECUST ON DIVECUST.`Customer_No`=DIVEORDS.`Customer_No` 
WHERE DIVEITEM.`Rental_Sale`='Rental' 
AND DIVECUST.`Name`='Louis Jazdzewski';
```
#### Output 

82

## Problem 5: What are the names of the wildlife that Mary Rioux might see on her trip. Are there any Shipwrecks there (give names)?

```sql
SELECT DISTINCT Common_Name FROM BIOLIFE 
INNER JOIN BIOSITE ON BIOLIFE.`Species_No` = BIOSITE.`Species_No` 
INNER JOIN SITES ON BIOSITE.`Site_No`=SITES.`Site_No` 
INNER JOIN DEST ON SITES.`Destination_No`=DEST.`Destination_No`
INNER JOIN DIVEORDS ON DEST.`Destination_Name`=DIVEORDS.`Destination`
INNER JOIN DIVECUST ON DIVEORDS.`Customer_No`=DIVECUST.`Customer_No`
WHERE DIVECUST.`Name`='Mary Rioux'
```
### Output：
Common_Name: Swell Shark, Bat Ray, California Moray, Lingcod, Cabezon, Yellow Jack, Redtail Surfperch, White Sea Bass, Rock Greenling, Senorita, Surf Smelt

## Problem 6: Who is paying the maximum amount for single type of rental equipment?

```sql
SELECT i.Qty * s.Rental_Price as Total,c.Name
FROM DIVEITEM i inner join DIVEORDS o ON i.`Order_No`=o.`Order_No` 
INNER JOIN DIVECUST c on c.`Customer_No`=o.`Customer_No`
INNER JOIN DIVESTOK s on i.Item_No = s.Item_No WHERE i.`Rental_Sale`='Rental' 
ORDER BY Total desc LIMIT 1;
```
### Output:
Total: 150   
Name: Sven Schiro


## Problem 7: What are the names of customers who are paying in cash?

```sql
SELECT Name FROM DIVECUST INNER JOIN DIVEORDS 
ON DIVECUST.`Customer_No`=DIVEORDS.`Customer_No` 
WHERE DIVEORDS.`PaymentMethod`='Cash'
```

### Output:
Name: Mary Rioux, Keith Lucas, Marcus Selby, Tony Rees, Lorraine Vega

## Problem 8: Produce a list of all equipment being rented for a dive vacation that costs more than $30000, make the list in descending order of the rental price of the equipment.

```sql
SELECT s.`Description`,s.`Rental_Price`,i.`Rental_Sale`,o.`VacationCost` 
FROM DIVESTOK s INNER JOIN DIVEITEM i ON s.`Item_No`=i.`Item_No` 
INNER JOIN DIVEORDS o ON i.`Order_No`=o.`Order_No` WHERE i.`Rental_Sale`='Rental' 
AND o.`VacationCost`>30000 
ORDER BY s.`Rental_Price` desc
```

### Output:
```{r,echo=FALSE}
q7 <- read.csv("q7.csv")
library(knitr)
library(pander)
pandoc.table(q7)
```

## Problem 9: Who is renting teal colored equipment?

```sql
SELECT DISTINCT Name FROM DIVECUST c INNER JOIN DIVEORDS o 
ON c.`Customer_No`=o.`Customer_No` INNER JOIN DIVEITEM i 
ON o.`Order_No`=i.`Order_No` INNER JOIN DIVESTOK s on i.`Item_No`=s.`Item_No` 
WHERE i.`Rental_Sale`='Rental' AND s.`Description` LIKE '%Teal'
```
### Output:
Name: David Burgett, Ken Soule

## Problem 10: Which locations have an avg temperature of more than 75 degrees Farenheit and a travel cost of under $4000?

```sql
SELECT d.`Destination_Name` FROM DEST d 
WHERE d.`Avg_Temp_F`>75 AND d.`Travel_Cost`<4000 
```

### Outcome:
Destination_Name: Cozumel, Florida


## Problem 11: Who spend(s) the most amount of money on traveling cost?

```sql
SELECT c.Name, d.`Travel_Cost` FROM DIVECUST c INNER JOIN DIVEORDS o 
ON c.`Customer_No`=o.`Customer_No` INNER JOIN DIVEITEM i 
ON o.`Order_No`=i.`Order_No` INNER JOIN DIVESTOK s 
ON i.`Item_No`=s.`Item_No` INNER JOIN DEST d 
ON d.`Destination_Name`=o.`Destination` 
ORDER BY d.`Travel_Cost` desc LIMIT 1
```
### Output:
Name: Marcus Selby

Travel_Cost: 5000
 
